Musical Instrument Classification Using a Hybrid Neural Network

### 2021-08-13 JMMPaper/PaperDrafts/main.tex
Updates to Introduction Section

Digital audio analysis and classification has become a very active field in the last few years. The exact nature of each task can differ drastically from archival management, to security, or to commercial usage. In each case, we seek to place audio files into distinct categories called \textit{classes} based on inherent properties of the data \cite{James,Khan,Liu}. Consider a collection of audio files, each containing a single note performed by some musical instrument and the task of matching the contents of the file to the instrument that it most closely resembles. This type of classification task is near trivial for humans, but given the terabytes and petabytes of audio data in the modern world, it is impractical at a large scale. However, a computer has no difficulty processing large volumes of raw information, but encoding an explicit and accurate instruction set for classification is unreasonably challenging. For this reason, we turn to a neural network to combine the computational efficiency of a computer, with the decision-making architecture of a simulated brain \cite{Geron}.

At a basic level, a neural network is a machine learning model that allows a set of inputs, $x$, called \textit{features}, to be transformed into a set of outputs, $y$,  called \textit{predictions} using a set of \textit{parameters}, $\Theta$ \cite{Geron,Goodfellow,Virtanen}. Audio classification with machine learning is also a well-explored field, with much research in developing an appropriate set of features, also called \textit{predictors}, for input \cite{James,Liu,Mierswa,Zhang}. Since the each classification task or data set may differ greatly from any other, each model and data set often requires a unique combination of features which allows for the best possible performance \cite{Virtanen}. For musical instrument classification we have chosen predictors that represent the contents of an audio file using two different \textit{modalities}. We call this type of classification \textit{multimodal learning} \cite{Ngiam}.

Multimodal learning differs from other types of supervised learning in that a model is designed to accept and process multiple inputs simultaneously that are each different representations of the same data sample. Data sets that include or can be transformed into audio + visual, audio + text, or even text + text information, are all examples of multimodal data sets \cite{Li}. For digital audio classification, we have chosen to represent the contents of an audio file by decomposing it into an \textit{image + vector} information format. We do this by constructing a 2D spectrogram and a 1D vector of features from the same waveform. Data from these different modes encodes complementary information which allows a model to develop parameters that can more thoroughly classify a sample when compared to unimodal learning methods \cite{Li}

To accommodate for this multimodal input, we have constructed a \textit{hybrid neural network} (HNN) that utilizes two input branches, each with its own set of layers to handle an input mode. The spectrogram image is processed with a Convolutional Neural Network (CNN) which uses layers of 2D convolution and 2D pooling to generate a feature map, which is then flattened and transformed by repeated dense layers - this is visualized on the left side of Fig. (\ref{fig:Architecture}). The feature vector input is processed by a Multilayer Perceptron (MLP) which uses repeated dense layers to transform the input, and is visualized on the right side of Fig. (\ref{fig:Architecture}). The output of each branch is concatenated into a single layer, which is further transformed into a single output prediction found at the bottom on Fig. (\ref{fig:Architecture}). This means that the neural network learns a set of parameters which allows for mapping of \textit{both} input modes into a single prediction. Since the properties of the input features are critically important to classification success, multi-representation learning shows a great deal of promise and widespread applicability \cite{Khan,Li,Liu,Virtanen}.

@@

A neural network defines a type of machine learning algorithm that is inspired from the human brain \cite{Geron,Goodfellow}. Where biological brains are constructed from neurons and connected through axons, neural networks are constructed from artificial neurons and connected through transformation functions. The numerical value contained within an artificial neuron is the \textit{activation} of that neuron \cite{Levine}. Neurons are organized into groups called \textit{layers}, which interact with other layers through mathematical transformations such as matrix multiplications or convolutions \cite{Goodfellow}. Because of this, each layer $l$ in a network can be considered to as function, $f^{(l)}$ that accepts an array of activations from the previous layer $x^{(l-1)}$ and uses a set of parameters to return an array of modified activations for the current layer, $x^{(l)}$. Thus each layer can be generally characterized by the expression:
\begin{equation}
    \label{eqn:LayerFunction}
    f^{(l)} : x^{(l-1)} \rightarrow x^{(l)}
\end{equation}

By connecting multiple layers in succession, we form a layer-chain structure that describes the flow of information in the neural network, and lets us model a neural network as a computational graph \cite{Geron,Goodfellow}. In a feed-forward network with $L$ layers, inputs $x^{(0)}$ and outputs $y^*$, we can represent a layer-chain such as:
\begin{equation}
    \label{eqn:LayerChain}
    x^{(0)} \rightarrow f^{(0)} \rightarrow f^{(1)} \rightarrow ... \rightarrow f^{(L-2)} \rightarrow f^{(L-1)} \rightarrow y^*
\end{equation}
The structure of a neural network's layer-chain defines it's \textit{architecture}. The choice of architecture indicates the \textit{hypothesis-space} of the model a determines the set of all possible solutions that a trained network can attain \cite{Goodfellow,James}. For deep neural networks this function chain can be dozens of layers long, and contain upwards of millions of parameters. Increasingly complicated networks allow for a high dimensional solution spaces, which potentially enables the model to describe more complicated systems.

>>>>

Digital audio analysis and classification has become a very active field in the last few years. The exact nature of each task can differ drastically from archival management, to security, or to commercial usage. In each case, we seek to place audio files into distinct categories called \textit{classes} based on inherent properties of the data \cite{James,Khan,Liu}. Consider a collection of audio files, each containing the sound of a single note performed by some musical instrument and the task of matching the categorizing the file based on what instrument the contents most closely resembles. This type of classification task is near trivial for humans but given the terabytes and petabytes of audio data in the modern world, it is impractical at a large scale. However a computer has no difficulty processing large volumes of raw information but encoding an explicit and reliable instruction set for classification is unreasonably challenging. For this reason, we turn to a neural network to combine the computational efficiency of a computer, with the decision-making architecture of a simulated brain \cite{Geron}.

At a basic level, a neural network is a machine learning model that allows a collection of inputs, $x$, called \textit{features}, to be transformed into a collection of outputs, $y$,  called \textit{predictions} through a set of adjustable values, $\Theta$, called \textit{parameters} \cite{Geron,Goodfellow,Virtanen}. In machine learning audio classification, much research is devoted to engineering and developing an appropriate ensemble of features that can be used to improve classification ability \cite{James,Liu,Mierswa,Zhang}. Since the nature of classification tasks and audio datasets may differ greatly from any other - each task must often develop a unique model and combination of features that will allow for the best possible performance \cite{Virtanen}. For musical instrument classification we have produced features that will represent the contents of each audio file using two different \textit{modalities}. Consequently, we must craft the neural network structure so that it may simultaneously accommodate these different input forms. We call this type of classification \textit{multimodal learning} \cite{Ngiam}.

Multimodal learning differs from other types of supervised learning in that a model is designed to accept and process multiple inputs forms that are each different representations of the same data sample, at the same time. Data sets that include or can be transformed into multiple forms such as \textit{audio + visual}, \textit{audio + text}, or even \textit{text + text} information, are all examples of multimodal data sets \cite{Li}. For digital audio classification, we have chosen to represent the contents of an audio file by decomposing it into an \textit{image + vector} information format. We do this by constructing a 2-dimensional spectrogram and a 1-dimensional vector of features from the same audio file. Data from these different modes encodes complementary information which allows a model to develop parameters that can more thoroughly classify a sample when compared to singlemodal learning methods \cite{Li}.

To account for this multimodal input, we have constructed a \textit{hybrid neural network} (HNN) that utilizes two different input layers where each receives its prescribes input format and is handled by an appropriate layer chain. The 2D spectrogram image is processed with a Convolutional Neural Network (CNN) which uses layers of 2D convolution and 2D pooling to generate a feature map, is then flattened and transformed by repeated dense layers. This behavior is visualized on the left side of Fig. (\ref{fig:Architecture}). The 1D feature vector is processed by a Multilayer Perceptron (MLP) which uses repeated dense layers to transform the input, and is visualized on the right side of Fig. (\ref{fig:Architecture}). The output of each branch is concatenated into a single dense layer, which is further transformed into a single output prediction found at the bottom on Fig. (\ref{fig:Architecture}). This means that the neural network learns a set of parameters which allows for mapping of \textit{both} input modes into a single prediction. Since the properties of the input features are critically important to classification success, multi-representation learning shows a great deal of promise and widespread applicability \cite{Khan,Li,Liu,Virtanen}.

@@

A neural network defines a type of machine learning algorithm that is inspired from the human brain \cite{Geron,Goodfellow}. Where biological brains are constructed from neurons and connected through axons, neural networks are constructed from artificial neurons and connected through transformation functions. The numerical value contained within an artificial neuron is the \textit{activation} of that neuron \cite{Levine}. Neurons are organized into groups called \textit{layers}, which interact with other nuurons in other layers through mathematical transformations such as matrix multiplications or convolutions \cite{Goodfellow}. Because of this interaction, each layer $(l)$ in a network can be considered as function, $f^{(l)}$ that accepts an array (vector,matrix,tensor,etc.) of activations from the \textit{previous} layer $x^{(l-1)}$ and uses a set of \textit{parameters} to return an array of modified activations for the \textit{current} layer, $x^{(l)}$. Thus each layer can be generally characterized by the expression:
\begin{equation}
    \label{eqn:LayerFunction}
    f^{(l)} : x^{(l-1)} \rightarrow x^{(l)}
\end{equation}

By connecting multiple layers in succession, we form a chain-like structure of layers that describes the flow of information in the neural network. This allows us to model a neural network as a directed computational graph where each node is a layer that performs an operation  \cite{Geron,Goodfellow}. In a feed-forward network with $L$ layers, inputs $x^{(0)}$ and outputs $x^{(L-1)} = y^*$, we can represent a layer-chain such as:
\begin{equation}
    \label{eqn:LayerChain}
    x^{(0)} \rightarrow f^{(0)} \rightarrow f^{(1)} \rightarrow ... \rightarrow f^{(L-2)} \rightarrow f^{(L-1)} \rightarrow y^*
\end{equation}
The structure of a neural network's layer-chain defines it's \textit{architecture}. The choice of architecture indicates the \textit{hypothesis-space} of the model a determines the set of all possible solutions that a trained network can attain \cite{Goodfellow,James}. For deep neural networks this function chain can be dozens of layers long, and contain upwards of millions of parameters. Increasingly complicated networks allow for a high dimensional solution spaces, which potentially enables the model to describe more complicated systems.
